%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Common
%
\newglossaryentry{controller}
{
  name=Domain Controller,
  description={Represents a running application that holds a database of resources and 
    reservations. Users can access and modify the database through the controller user 
    interface(s). The controller runs a scheduler that process the database.}
}
\newglossaryentry{connector}
{
  name=Connector,
  description={\TODO{}}
}
\newglossaryentry{user}
{
  name=User,
  description={\TODO{}}
}
\newglossaryentry{person}
{
  name=Person,
  description={\TODO{}}
}
\newglossaryentry{date-time-slot}
{
  name=Date/Time Slot,
  description={Represents a time between specific starting and ending moment. Can be 
    specified by a starting date and time and a duration.}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Resources
%
\newglossaryentry{resource}
{
  name=Resource,
  description={An entity that can be scheduled by a scheduler.}
}
\newglossaryentry{device}
{
  name=Device,
  description={A special type of resource that represents a videoconference device (e.g., 
    terminal, gateway or multipoint control unit).}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Reservation
%
\newglossaryentry{reservation-request}
{
  name=Reservation Request,
  description={A request made by an user to book given resources for a specific date/time 
    slot(s). Reservation request can be \textbf{incomplete} or \textbf{complete}. Incomplete 
    request must be filled in by additional information to become complete.
    \TODO{Do we break the request also to meetings to be able to schedule complete meeting 
    request that is part of not complete reservation request?}}
}
\newglossaryentry{reservation-meeting}
{
  name=Reservation Meeting,
  description={Represents allocated resources for a single date/time slot based on a 
  complete reservation request.}
}
\newglossaryentry{reservation}
{
  name=Reservation,
  description={Represents a partially or fully scheduled complete reservation request. 
  Contains one or more reservation meetings.}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Scheduler
%
\newglossaryentry{scheduler}
{
  name=Scheduler,
  description={Scheduler is a component of a domain controller that process complete 
    reservation requests and schedule them to reservations.}
}