\section{Example 1}

This example describes the simplest scenario in which the resource database contains only one MCU that support H.323 and SIP. Some simple reservation requests are listed.

\subsection{Resource Database}

Resource database is composed of the following resources:
\begin{itemize}

\item Multipoint device (\emph{mcu}) that supports H.323 and SIP 
  technologies.
  
\begin{EntityExample}{DeviceResource}{mcu}{Multipoint device for H.323 and SIP}
technologies: [H323, SIP], 
ipAddress: 147.251.99.3,
capabilities: [
  ReceiveCapability, SendCapability,
  VirtualRoomsCapability,
  SignalingClientCapability {technologies: [H323], server: proxy},
  SignalingClientCapability {technologies: [SIP], server: proxy}
]
\end{EntityExample}

\end{itemize}

\subsection{Reservation Requests}

\paragraph{Request 1}
\begin{enumerate}
\item An user creates a reservation request.

\begin{EntityExample}{ReservationRequest}{reservationRequest}{}
purpose: Science,
requestedSlots: [DateTimeSlot(2012-05-30T14:00, P1H)],
requestedResources: [
  ExternalEndpointSpecification {
    technologies: [H323],
    count: 5,
    persons: [
      PersonByIdentity(srom@cesnet.cz),
      PersonByIdentity(hopet@cesnet.cz),
      PersonByIdentity(janru@cesnet.cz)
    ]
  }
]
\end{EntityExample}

\item A domain controller stores the reservation request and based on it the controller creates a compartment request which is also stored.

\begin{EntityExample}{CompartmentRequest}{compartmentRequest}{}
purpose: Science,
requestedSlot: AbsoluteDateTimeSlot(2012-05-30T14:00, P1H),
requestedResources: [
  ExternalEndpointSpecification {
    technologies: [H323],
    count: 5,
    persons: [
      PersonByIdentity(srom@cesnet.cz),
      PersonByIdentity(hopet@cesnet.cz),
      PersonByIdentity(janru@cesnet.cz)
    ]
  }
]
\end{EntityExample}

\item The compartment request is incomplete because all requested persons must
confirm theirs attendance. After all persons are confirmed or rejected
the request become complete.

\item Scheduler watches for complete compartment requests, so the complete 
request is now processed by the scheduler. The scheduler knows that he must 
interconnect 5 unknown H.323 endpoints and thus he searches his resource 
database and finds the \emph{mcu}. He must setup a virtual room in the 
\emph{mcu} for the specified date/time and five participants.

\item The scheduler creates allocated compartment entity to which he appends 
allocated virtual room resource. The scheduler also creates a reservation
entity to which he appends the allocated compartment entity.

\item The reservation is propagated to the reservation request owner.

\TODO{Complete the example} 

\end{enumerate}
